Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}

### 2023-05-18 paper/paper.tex
#15 add introduction with business metaphor

\begin{abstract}
I use historical data on construction activity and territorial history to estimate the effect of switching between territories on urban construction in the Holy Roman Empire. I exploit variation from dynastic extinctions, which caused cities to become part of different lineages' territorial holdings. I find that an X\% increase in the number of cities possessed by the ruling family is associated with a Y\% increase in construction activity.
\end{abstract}

@@

\section{Research Question}

When a ruling dynasty goes extinct, what happens to their lands?

\subsection{Disambiguation}
Let me be more precise. Concepts to be defined:
\begin{itemize}
    \item[ruling dynasty] In the context of the HRE. Until the 19th century geopolitics was not done in terms of nation-states, but in terms of the territorial holdings of noble families.
    \item[their lands] The cities in that dynasty's possession at the time of its extinction.
    \item[what happens] It's complicated and precisely this is what I need to clarify first.
\end{itemize}

@@

@@

>>>>

\begin{abstract}
When a ruling dynasty goes extinct, what happens to their lands? I use historical data on construction activity and territorial history to estimate the effect of switching between territories on urban construction in the Holy Roman Empire. I exploit variation from dynastic extinctions, which caused cities to become part of different lineages' territorial holdings. I find that an X\% increase in the number of cities possessed by the ruling family is associated with a Y\% increase in construction activity.
\end{abstract}

@@

\section{Introduction}

% one paragraph to get the reader from where he stands, to understanding and caring about the research question.
The geopolitics of central Europe, throughout its history and until very recently, have been defined by noble families struggling over control of a vast array of fragmented territories - unlike, for example, China. Rulers have played an important role in the fates of their nations. Examples come to mind. But did such geopolitical drama have real effects on the lives of people?

Let us consider Example city A. From 1XXX on, A belonged to the dukedom of Saxe-Weimar-Eisenach, until its rulers, the House of Ascania, went extinct in 1XYZ. The territory of Saxe-Weimar-Eisenach was then subsumed into the much larger Electorate of Saxony, ruled by the Albertine branch of the House of Wettin. By the same token, A changed hands from the Ascanians to the Albertines.

What happens on the political stage is clear; the economic ("real") implications are not. In the language of business acquisitions: do such extinction-and-succession events manifest similar to a strategic acquisition, with some degree of reorientation? Do the participants benefit from synergies and reduced transaction costs? Or are these events more akin to mere ownership transfers?

\subsection{Disambiguation}
Let me be more precise. Concepts to be defined:
\begin{itemize}
    \item[ruling dynasty] In the context of the HRE. Until the 19th century geopolitics was not done in terms of nation-states, but in terms of the territorial holdings of noble families. Leadership changes can be split into within-family and across-family - we care about the latter.
    \item[their lands] The cities in that dynasty's possession at the time of its extinction.
    \item[what happens] It's complicated and precisely this is what I need to clarify first.
\end{itemize}

@@

The effects of different types of rule.

Leaning on Dube and Harish (2022): family-individual characteristics vs family-leadership characteristics (Habsburgs may be docile by nature but make aggressive decisions in the interest of their realm out of geopolitical necessity)

@@

Do mention that, because construction is a rare event, the time series for individual cities aren't as meaningful as averages.
